\documentclass[../main.tex]{subfiles}

\graphicspath{{\subfix{../images/}}}
\begin{document}

\chapter{Data Modeling}
\section{Introduction to Data Modeling}
Data Modeling is process of defining how data is stored and the relationships and the relationships that exists among different entities in your data. \\ 
We refer to Organization of data inside a database as a \textbf{Schema}. \\ 
To develop schema think about your database and you need to ask these questions:
\begin{itemize}
	\item{What does my application do?}
	\item{What data will I store?}
	\item{How will users access this data?}
	\item{What data will be most valuable to me?}
\end{itemize}

Asking these questions will help you to:
\begin{itemize}
	\item{descibe your task as well those of users. }
	\item{What you data looks like?}
	\item{the relationships among the data.}
	\item{The tooling you plan to have.}
	\item{The access patterns that might emerge.}
\end{itemize}

\medskip
A good data model can: 
\begin{itemize}
	\item{Make it easier to manage data}
	\item{Make queries more efficient}
	\item{Use less memory and \gls{CPU}}
	\item{Reduce cost}
\end{itemize}

As a general pricipal of MongoDB data that is accessed together should be stored together. \\ 
Collections do not enforce any document structure by default. That means document even those in the same collection can have different structures. \\
Each document at MongoDB can be different which is known as \textbf{Polymorphism}. \\ 
Embedded document model enables us to build complex relationships among data. You can normalize your data by using database references. \\ 
The key pricipal here is how your application will use the data rather than how it's stored in the database.



\section{Types of Data Relationships}
\subsection{Different types of relationships data can have: }
\begin{itemize}
	\item{One-to-one}
	\item{One-to-many}
	\item{Many-to-Many}
\end{itemize}

\subsubsection{One-to-One}
A relationship where a data entity in one set is connected to exactly one data entity in another set.

\subsubsection{One-to-Many}
A relationship where a data entity in one set is connected to any number of entities in another set.
\subsubsection{Many-to-Many}
A relationship where any number of data entities in one set are connected to any number of data entities in another set.


\subsection{Two main ways to model these relationships: }
\begin{itemize}
	\item{Embeding}
	\item{Referencing}
\end{itemize}
\subsubsection{Embeding}
We take related data and insert it into our document.
\subsubsection{Referencing}
When we refer to documents in another collection in our document. Reference is made by their respective \lstinline{ObjectID}.

\section{Modeling Data Relationships}
\subsection{Embedding Data in Documents}
Also known as nested document.
\begin{itemize}
	\item Avoids application joins. 
	\item Provides better performance for read operatinos.
	\item Allows developers to update related data in a single write operations.
\end{itemize}
\subsubsection {Issues:}

\begin{itemize}
	\item Embedding data into a single document can create large documents.
	\item{Large documents have to be read into memory in full, which can result in a slow application performance for you end user.}
	\item{Continuesly adding data without limit creates \textbf{unbounded documents.}}
	\item{Unbounded Documents may exceed the \gls{BSON} document threshold of 16MB. This is schema anit-patterns which you should avoid.}
\end{itemize}

\subsection{Referencing Data in Documents}
When we want to store data in two different collections but also ensure that those collections are related, we can use \textbf{References}.
References save the $_id$ field of one document in another document as a link between the two.
They are simple and sufficient for most use cases. Using references is called \textbf{linking} or \textbf{data normalization}.
It avoid duplication of data and in most cases smaller documents, however referencing data you'll need to query from multiple documents costing extyra resources and impacts read performance. 

\bigskip

\begin{tabular}{|c|c|}
\toprule
\textbf{Embedding} & \textbf{Referencing}  \\ 
\midrule
$\checkmark$ Single query to retrieve data  & $\checkmark$ No duplication   \\ \hline
$\checkmark$ Single operation to update/delete data  & $\checkmark$ Smaller Documents   \\ \hline
! Data Duplication  & ! Need to join data from multiple documents  \\ \hline
! Large documents  &  \\ 
\bottomrule
\end{tabular}

\section{Scaling a Data Model}
Optimum efficiency of: 
\begin{itemize}
	\item query result depends on 
	\item memory usage
	\item \gls{CPU} usage 
	\item storage
\end{itemize}

\begin{itemize}
	\item When creating a data model, avoid documents that are unbounded. Document size grows infinitely.
		Problems as teh array grows larger in embedded document.
		\begin{itemize}
			\item It will take up more space in memory 
			\item may impact write performance as entire document is rewritten into MongoDB data storage.
			\item it will be difficult to perform pagination of that array.
			\item Maximum document size of 16MB will lead to storage problems.
		\end{itemize}
	\item benefit is that we can retrieve all document in single read.
	\item Avoid: 
		\begin{itemize}
			\item More than the document size limit of 16 MB.\@
			\item Poor query performance.
			\item Poor write performance.
			\item Too much memory being used.
		\end{itemize}
\end{itemize}

\section{Using Atlas Tools for Schema Help}
Schema design patterns are guidelines that help developers plan, organize, and model data.
Schema anti-patterns result in suboptimal performance and non scalable solutions. Eg: 
\begin{itemize}
	\item massive arrays.
	\item Massive number of collections.
	\item Bloated documents. 
	\item Unnecessary indexes.
	\item Queries without indexes.
	\item Data that's accessed together, but stored in different collections.
\end{itemize}

Tools in Atlas are \textbf{Data Explorer} and \textbf{Performance Advisor}


\printglossaries

\end{document}

